Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-26 sections/compiler.tex
add dimensionalization

\subsection{Concordization}

>>>>

    However, it is cumbersome to write the dimensions
    in loop programs, and most tensor compilers have a means of specifying the
    dimensions automatically. We refer to the process of discovering suitable
    dimensions as \textbf{dimensionalization}. \willow{Teo, please explain
    alternative approaches}>

    In finch, we use a straightforward algorithm to automatically determine
    suitable dimensions for loops and declaration statements (output tensors).
    Finch determines the dimension of a loop index i from all of the tensors
    using i in an access, as well as the bounds in the loop itself, and operates
    similarly for declarations.

    Our algorithm uses the following principles, assuming dimension types form a
    lattice:
    \begin{enumerate}
        \item We assign dimensions to indices.
        \item Using an index in an access “hints” that the index should have the corresponding dimension.
        \item Loop dimensions are equal to the “meet” of all hints in the loop body
        and any existing dimensions in the loop bounds. The meet usually asserts
        that dimensions match, but may also e.g. propagate info about parallelization
        \item The \mintinline{julia}{_} symbol represents a dimensionless quantity at the bottom of the dimension lattice.
        \item We assign dimensions to declarations.
        \item Left hand side (updating) tensor access “hint” the size of that tensor
        \item The dimensions of a declaration are the “meet” of all hints from
        the declaration to the first read.
        \item The new dimensions of the declared tensor are used when the tensor is on the right hand side (reading)
        access.
    \end{enumerate}

    For example, consider the following code

    \begin{minted}{julia}
    A = fsprand(3, 4, 0.5)
    B = fsprand(4, 5, 0.5)
    C = Tensor(Dense(SparseList(Element(0.0))))
    @finch begin
    C .= 0
    for i = 1:3
        for j = _
            for k = _
                C[i, j] += A[i, k] * B[k, j]
            end
        end
    end
    \end{minted}

    In the above code, the second dimension of A must match the first dimension of B. Also, the first dimension of A must match the i loop dimension, 1:3. Finch will also resize declared tensors to match indices used in writes, so C is resized to (1:3, 1:5). If no dimensions are specified elsewhere, then Finch will use the dimension of the declared tensor.

    Dimensionalization occurs after wrapper arrays are de-sugared. You can
    therefore exempt a tensor from dimensionalization by wrapping the
    corresponding index in ~ to produce a "PermissiveArray". For example,

    \begin{minted}{julia}
    @finch begin
    y .= 0
    for i = 1:3
        y[~i] += x[i]
    end
    \end{minted}

    does not set the dimension of y, and y does not participate in dimensionalization.

\subsection{Concordization}
    Finch uses
